---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r load_data}
data <- read.csv("activity.csv")
data$date_d <- as.Date.factor(data[[2]])
library(ggplot2)
data_a <- aggregate(.~date,data[1:2], FUN=sum, na.rm=TRUE)
data_b <- aggregate(.~interval,data[c(1,3)], FUN=mean, na.rm=TRUE)
```


## What is mean total number of steps taken per day?
```{r Mean_total_per_day}
ggplot(data_a, aes(steps)) + geom_histogram()  + labs(title = "Distribution of steps count per day")
mean1 <- as.character(floor(mean(data_a$steps)))
median1 <- as.character(median(data_a$steps))

```


The mean of total number of steps taken per day is `r mean1`, and de median is `r median1`.


## What is the average daily activity pattern?
```{r Mean_per_interval}
ggplot(data_b, aes(interval,steps)) + geom_line() + labs(title = "Average steps per interval")
max_interval <- as.character(data_b[data_b$steps == max(data_b$steps),1])
max_steps <- as.character(floor(data_b[data_b$steps == max(data_b$steps),2]))
```


The 5-minute interval with the maximum average of steps is the interval `r max_interval` with `r max_steps` steps.


## Imputing missing values
```{r}
nas <- sum(is.na(data[1]))
```
There are `r nas` missing values in the dataset.

Now we create a new dataset filling the missing values with te average per interval.
```{r Mean_total_per_day_missing_filled}
data2 <- merge(data, data_b, by.x = "interval", by.y = "interval")
data2[is.na(data2[2]),2] <- data2[is.na(data2[2]),5]
data2_a <- aggregate(.~date,data2[2:3], FUN=sum)
ggplot(data2_a, aes(steps.x)) + geom_histogram()  + labs(title = "Distribution of steps count per day")
mean2 <-as.character(floor(mean(data2_a$steps.x)))
median2 <- as.character(floor(median(data2_a$steps.x)))
```

The new mean of total steps per day is `r mean2`, and the median is `r median2`. Filling the mission values the mean and the median are now equals.


## Are there differences in activity patterns between weekdays and weekends?
We generate a new variable that indicates if is a weekday or a weekend.
```{r Mean_per_interval_Weekdays}
data2$day <- ifelse(weekdays(data2$date_d)=="domingo" | weekdays(data2$date_d)=="sábado","weekend","weekday")

data2_c <- aggregate(.~interval+day,data2[c(1,2,6)], FUN=mean)
ggplot(data2_c, aes(interval, steps.x)) + geom_line() + facet_grid(.~day) + labs(title="Average steps per interval") + labs(y="Average steps")
```